---
title: "Hacking for Historians"
subtitle: "An Introduction to Digital History with R"
author: "Matthias Bannert** (@whatsgoodio) and Ina Serif++"
institute: "ETH Zurich**, Uni Basel++"
date: "2019-09-25"
output:
  xaringan::moon_reader:
    css: ["robot-fonts","more.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---

##  teachers:~$ whoami

.pull-left[

### Matthias Bannert ([@whatsgoodio](https://twitter.com/whatsgoodio))

- Data DevOps @KOF, ETH Zurich 
- PhD in Economics 
- 20+ years programming experience
 
]

.pull-right[

### Ina Serif

]

**course material**: [github.com/mbannert/dh-2019](https://github.com/mbannert/dh-2019)



---

## Who Are You?

---

class: center, middle

## Why Would a Historian Want to Program? 

<!--
Show demos
1. API examples
2. visualization 
3. Text Mining & Regular Expressions
-->


---


## Why Would a Historian Want to Program? 


- Access to Data Sources (**A**pplication **P**rogramming **I**nterfaces) 



---


## Why Would a Historian Want to Program? 


- Access to Data Sources (**A**pplication **P**rogramming **I**nterfaces) 

- Visualization


---


## Why Would a Historian Want to Program? 


- Access to Data Sources (**A**pplication **P**rogramming **I**nterfaces) 

- Visualization

- Text Mining & Regular Expressions

---

## Why Would a Historian Want to Program? 


- Access to Data Sources (**A**pplication **P**rogramming **I**nterfaces) 

- Visualization

- Text Mining & Regular Expressions

- Reproducibility



---


## Why Would a Historian Want to Program? 


- Access to Data Sources (**A**pplication **P**rogramming **I**nterfaces) 

- Visualization

- Text Mining & Regular Expressions

- Reproducibility

- Code as Communication Channel

- ...


---

## Background Poll


```{r,echo=FALSE}
library(grid)
library(shiny)
source("../R/bullet.R")
fluidPage(sidebarLayout(sidebarPanel(
  numericInput("members", "Number of participants", 20),
  numericInput("rstats", "R Statistical langauge", 8),  
  numericInput("html", "HTML/CSS/JS devs", 8),
  numericInput("shiny", "shiny framework", 8)
  ),
  mainPanel(
    renderPlot({
      techavg <- mean(c(input$rstats,
                        input$html,
                        input$shiny) / input$members)*100
  df1 <- data.frame(units = c("R(%)","WebDev(%)","shiny(%)"),
                    low = c(25,25,25),
                    mean = c(50,50,50),
                    high = c(100,100,100),
                    target = c(techavg,techavg,techavg),
                    value = c(100*(input$rstats/input$members),
                              100*(input$html/input$members),
                              100*(input$shiny/input$members)
                              )
                    )
  g <- gridBulletGraphH(df1,bcol = c("#999999","#CCCCCC","#E1E1E1"),
                   vcol = "#333333",font = 20)
  g + title(paste("Usage of Technologies Among Participants", sep=" "))
})  
    )
  ))
```

---

## Several introduction slides here... 

---

## Some tasks




