---
title: "Hacking for Historians"
subtitle: "Programming with Data"
author: "Matthias Bannert** (@whatsgoodio) and Ina Serif++"
institute: "ETH Zurich**, Uni Basel++"
date: "2019-09-25"
output:
  xaringan::moon_reader:
    css: ["robot-fonts","more.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


## Exploring Built-In Data 

Show all dataset available from currently loaded packages.
```{r, eval=FALSE}
data()
```

---


## Exploring Built-In Data 

Show all dataset available from currently loaded packages.
```{r, eval=FALSE}
data()
```

Load a specific dataset from the library *historydata*

```{r}
library(historydata)
data("catholic_dioceses")
```

---



## Exploring Built-In Data 

Show all dataset available from currently loaded packages.
```{r, eval=FALSE}
data()
```


Load a specific dataset from the library *historydata*
```{r}
library(historydata)
data("catholic_dioceses")
```


See what's in memory. <br> 
Note that there can be multiple datasets in memory at the same time!

```{r}
data("paulist_missions")
ls()
```

---

## Task 2:  Functions to Explore Dataset

Try out some functions to explore datasets from the 
*historydata*, *janeaustenr* and *fivethirtyeight*  packages or 
R's built-in datasets.

```{r, eval=FALSE}
str()
summary()
length()
nrow()
class()
pairs()
plot()
```

Does each function work for with each dataset? Is the output of, e.g., summary the same for all datasets?


## Writing Datasets to the File System


## Reading Data into R 






